\noindent
\large {\bf Organellar Genomes of White Spruce (Picea glauca): Assembly and Annotation} 


\normalsize 


\noindent \paragraph{Keywords:} gymnosperm, white spruce, organelle, genome assembly, sequencing, ABySS

\noindent \paragraph{Abstract:} 

The genome sequences of the plastid and mitochondrion of white spruce (Picea glauca) were
assembled from whole-genome shotgun sequencing data using ABySS. The sequencing data
contained reads from both the nuclear and organellar genomes, and reads of the organellar
genomes were abundant in the data as each cell harbors hundreds of mitochondria and plastids.
Hence, assembly of the 123-kb plastid and 5.9-Mb mitochondrial genomes were accomplished by
analyzing data sets primarily representing low coverage of the nuclear genome. The assembled
organellar genomes were annotated for their coding genes, ribosomal RNA, and transfer RNA.
Transcript abundances of the mitochondrial genes were quantified in three developmental tissues
and five mature tissues using data from RNA-seq experiments. C-to-U RNA editing was observed
in the majority of mitochondrial genes, and in four genes, editing events were noted to modify
ACG codons to create cryptic AUG start codons. The informatics methodology presented in
this study should prove useful to assemble organellar genomes of other plant species using
whole-genome shotgun sequencing data.

Chloroplast genomes of gymnosperms, including conifers, are well studied, but little is known
about the mitochondria of gymnosperms. In fact, only a single gymnosperm mitochondrion is
found in NCBI GenBank. This nearest related mitochondrial sequence is of the Prince Sago palm
(Cycas taitungensis) native to Taiwan, which diverged from the white spruce over a hundred
million years ago. No conifer mitochondrion genomes are to be found in GenBank at all, until
now.

Roughly one percent of the whole genome sequencing reads of white spruce are from its two
organellar genomes: the chloroplast and mitochondrion. We assembled these reads using ABySS
and found the mitochondrion genome to be nearly six megabases, which is unusually large for a
mitochondrial genome. Although many genes typical of mitochondria were found in the genome,
most open reading frames had no similarity to any known gene.

White spruce is an economically important species to the forestry industry of Canada. Insights
into the conifer mitochondrial genome will provide relevant new information to reconstruct the
evolution of this organelle genome relative to other plant lineages, and to identify which genes
of a conifer are uniquely inherited through the mitochondria. As the mitochondrial genome is
inherited maternally, and the plastid genome is inherited paternally, having a complete genome
sequence for both organelles would enable classifying the maternal and paternal species of hybrid
seed lots and determining the maternal and paternal lineage of saplings in breeding experiments.

\noindent \paragraph{Authors:} 

\paragraph{}

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
{
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|p{5cm}|l|l|}
\hline
\textbf{first name} & \textbf{last name} & \textbf{email}              & \textbf{country} & \textbf{organization}                                                          & \textbf{Web page}   & \textbf{corresponding?} \\ \hline
Shaun D             & Jackman            & sjackman@bcgsc.ca           & Canada           & BC Cancer Agency Genome Sciences Centre                                        & http://sjackman.ca/ & $\checkmark$                       \\ \hline
Rene                & Warren             & rwarren@bcgsc.ca            & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Ewan                & Gibb               & egibb@bcgsc.ca              & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     & $\checkmark$                       \\ \hline
Ben                 & Vandervalk         & benv@bcgsc.ca               & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Hamid               & Mohamadi           & hmohamadi@bcgsc.ca          & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Justin              & Chu                & cjustin@bcgsc.ca            & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Anthony             & Raymond            & traymond@bcgsc.ca           & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Stephen             & Pleasance          & spleasance@bcgsc.ca         & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Robin               & Coope              & rcoope@bcgsc.ca             & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Mark R              & Wildung            & wildung@wsu.edu             & United States    & School of Molecular Biosciences, Washington State University                   &                     &                         \\ \hline
Carol E             & Ritland            & critland@mail.ubc.ca        & Canada           & Department of Forest and Conservation Sciences, University of British Columbia &                     &                         \\ \hline
Jean                & Bousquet           & jean.bousquet@sbf.ulaval.ca & Canada           & Department of Forest and Environmental Genomics, Universite Laval              &                     &                         \\ \hline
Steven JM           & Jones              & sjones@bcgsc.ca             & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
Joerg               & Bohlmann           & bohlmann@mail.ubc.ca        & Canada           & Department of Forest and Conservation Sciences, University of British Columbia &                     & $\checkmark$                       \\ \hline
Inanc               & Birol              & ibirol@bcgsc.ca             & Canada           & BC Cancer Agency Genome Sciences Centre                                        &                     &                         \\ \hline
\end{tabular}%
}
}


