\noindent
\large {\bf A Hybrid Genome Assembler for Second- and Third-Generation Sequencing} 


\normalsize 


\noindent \paragraph{Keywords:} Assembly, Algorithms, Next-generation sequencing, Genome analysis

\noindent \paragraph{Abstract:} 

Genome assembly is challenged by the presence of large and complex repeats. The 2$^{nd}$
generation sequencing offers high throughput in low cost but the read length is inadequate to resolve 
large repeats. On the other hand, the 3$^{rd}$ generation sequencing can generate much longer reads for
spanning large repeats, but the error rate and sequencing cost are much higher.

This paper presented several new algorithms for assembling high-quality short reads and
low-quality long reads in two stages. In the correction stage, long reads are mapped onto an FM-index
 constructed from short reads and polished by FM-index extension. In the assembly stage, a
novel algorithm (called locality-sensitive backward search) is proposed to efficiently compute inexact
overlap among (partially) corrected long reads. The results indicated that the correction power and
accuracy is higher than existing methods. The developed assembler is able to assemble high-quality
genome (N50 in Mb) using low-coverage PacBio sequencing.

Availability: \url{https://github.com/ythuang0522/StriDe}


\noindent \paragraph{Authors:} 

\noindent \paragraph{} 

{
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{first name} & \textbf{last name} & \textbf{email}        & \textbf{country} & \textbf{organization}           & \textbf{corresponding?} \\ \hline
Yao-Ting            & Huang              & ythuang@cs.ccu.edu.tw & Taiwan           & National Chung Cheng University &  $\checkmark$                       \\ \hline
Ping-Ye             & Chen               &                       & Taiwan           & National Chung Cheng University &                         \\ \hline
\end{tabular}%
}
}
