\documentclass[11pt]{article}
\usepackage{epsfig}
\usepackage{amssymb,amsfonts,url}
\urlstyle{rm}   
\usepackage{color}
\usepackage[utf8]{inputenc}
\usepackage{wrapfig}
\usepackage{graphicx}
\newcommand{\junk}[1]{}
\usepackage{fullpage}

\date{}
\author{}

\pagenumbering{empty}

\begin{document}
\noindent
\large {\bf Rail-RNA: Scalable analysis of RNA-seq splicing and coverage} 


\normalsize 


\noindent \paragraph{Keywords:} Sequence alignment, RNA

\noindent \paragraph{Abstract:} 

RNA sequencing (RNA-seq) experiments now span hundreds to thousands of samples.
Current spliced alignment software is designed to analyze each sample separately. Consequently, no
information is gained from analyzing multiple samples together, and it requires extra work to obtain analysis
products that incorporate data from across samples.

 We describe Rail-RNA, a cloud-enabled spliced aligner that analyzes many samples at once.
Rail-RNA eliminates redundant work across samples, making it more efficient as samples are added.
For many samples, Rail-RNA is more accurate than annotation-assisted aligners. We use Rail-RNA to
align 667 RNA-seq samples from the GEUVADIS project on Amazon Web Services in under 16 hours
for US\$0.91 per sample. Rail-RNA outputs alignments in SAM/BAM format; but it also outputs (1) base-level 
coverage bigWigs for each sample; (2) coverage bigWigs encoding normalized mean and median
coverages at each base across samples analyzed; and (3) exon-exon splice junctions and indels (features)
in columnar formats that juxtapose coverages in samples in which a given feature is found. Supplementary
outputs are ready for use with downstream packages for reproducible statistical analysis. We use Rail-RNA
to identify expressed regions in the GEUVADIS samples and show that both annotated and unannotated
(novel) expressed regions exhibit consistent patterns of variation across populations and with respect to
known confounders.

Availability: Rail-RNA is open-source software available at \url{http://rail.bio}.

\noindent \paragraph{Authors:} 

\noindent \paragraph{} 

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
{
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{first name} & \textbf{last name}  & \textbf{email}     & \textbf{country}                                                  & \textbf{organization}                                                                                             & \textbf{corresponding?} \\ \hline
Abhinav             & Nellore             & anellore@gmail.com & United States                                                     & Johns Hopkins University                                                                                          &    $\checkmark$                     \\ \hline
Leonardo            & Collado-Torres      &                    & United States                                                     & Johns Hopkins University                                                                                          &                         \\ \hline
Andrew E.           & Jaffe               &                    & United States                                                     & Johns Hopkins University                                                                                          &                         \\ \hline
José                & Alquicira-Hernández &                    & \begin{tabular}[c]{@{}l@{}}United States\\ \& Mexico\end{tabular} & \begin{tabular}[c]{@{}l@{}}Johns Hopkins University\\ \& National Autonomous \\ University of Mexico\end{tabular} &                         \\ \hline
Christopher         & Wilks               &                    & United States                                                     & Johns Hopkins University                                                                                          &                         \\ \hline
Jacob               & Pritt               &                    & United States                                                     & Johns Hopkins University                                                                                          &                         \\ \hline
James               & Morton              &                    & United States                                                     & University of California San Diego                                                                                &                         \\ \hline
Jeffrey T.          & Leek                &                    & United States                                                     & Johns Hopkins University                                                                                          &                         \\ \hline
Ben                 & Langmead            & langmea@cs.jhu.edu & United States                                                     & Johns Hopkins University                                                                                          &         $\checkmark$                \\ \hline
\end{tabular}%
}
}

\end{document}
