\noindent
\large {\bf MuClone: Detection and classification of somatic mutations through probabilistic integration of clonal population structure} 


\normalsize 


\noindent \paragraph{Keywords:} Single nuclotide variants, Tumour clones, Mutation detection,
Mutation classification, Cancer evolution

\noindent \paragraph{Abstract:} 


Accurate detection and classification of somatic single
nucleotide variants (SNVs) is important in defining the clonal
composition of human cancers. Existing tools are prone to missing
low prevalence mutations and methods for classification of mutations
into clonal groups across the whole genome are underdeveloped.
Increasing interest in deciphering clonal population dynamics over
multiple samples in time or anatomic space from the same
patient is resulting in whole genome sequence (WGS) data from
phylogenetically related samples. We posited that injecting clonal
structure information into the inference of mutations from multiple
samples would improve mutation detection.

 We developed MuClone: a novel statistical framework for
simultaneous detection and classification of mutations across multiple
samples of a patient from whole genome or exome sequencing
data. The key advance lies in incorporating prior knowledge about
the cellular prevalences of clones to improve the performance
of detecting mutations, particularly low prevalence mutations. We
evaluated MuClone through synthetic and real data from spatially
sampled ovarian cancers. The results support the hypothesis that
clonal information improves the sensitivity without compromising the
specificity. In addition, MuClone classifies mutations across whole
genomes of multiple samples into biologically meaningful groups that
can provide additional phylogenetic insights and permits studying
clonal dynamics from WGS data.

Availability: MuClone is available from \url{http://compbio.bccrc.ca/}

\noindent \paragraph{Authors:} 

\noindent \paragraph{} 

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
{
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{first name} & \textbf{last name} & \textbf{email} & \textbf{country} & \textbf{organization}                                                                                                  & \textbf{corresponding?} \\ \hline
Fatemeh             & Dorri              &                & Canada           & University of British Columbia                                                                                         &                         \\ \hline
Sean                & Jewell             &                & United States    & University of Washington                                                                                               &                         \\ \hline
Alexandre           & Bouchard-Côtè      &                & Canada           & University of British Columbia                                                                                         &                         \\ \hline
Sohrab P.           & Shah               & sshah@bccrc.ca & Canada           & \begin{tabular}[c]{@{}l@{}}University of British Columbia\\ \& British Columbia Cancer \\ Research Center\end{tabular} &     $\checkmark$                    \\ \hline
\end{tabular}%
}
}
