\noindent
\large {\bf Consensus Representation Estimation of Lineage Expression \(CREoLE\) Algorithm for scRNA-seq} 


\normalsize 


\noindent \paragraph{Keywords:} Algorithms, Bioinformatics, Signal processing

\noindent \paragraph{Abstract:} 
Single-cell RNA-sequencing (scRNA-seq) is a promising technology widely used to
recapitulate gene expression trends through developmental progression of heterogeneous biological
tissue. Although several methods have sought to estimate pseudo-temporal gene expression trends, a
number of technical limitations presented by scRNA-seq remain, including high expression variability
and drop-out measurements, which complicate accurate trend estimation. Consensus Representation
Estimation of Lineage Expression (CREoLE) is an efficient and robust algorithm for automatically detecting
an underlying branching lineage structure and estimating smooth developmental trends of gene expression
by consensus without the need for model fitting or data filtering.

 Applied to synthetic data, CREoLE recapitulates underlying gene expression for each gene and
across each lineage with an average Pearson correlation coefficients of 0.983 $\pm$ 0.009. The impact of
simulated technical noise, drop-out measurements, and cell count reduction are evaluated. Our analysis
suggests that Creole is robust to additive noise and smaller initial cell populations (as low as 25\%
initial population). CREoLE correlates with synthetic expression trends with a mean Pearson’s correlation
coefficient above 0.9 in all cases. Applied to real data, CREoLE accurately identifies lineage structure and
computes high-resolution consensus trends that align closely with published findings.

Availability: CREoLE is free and open-source software available from \url{https://github.com/schaugf/creole}


\noindent \paragraph{Authors:} 

\noindent \paragraph{} 

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
{
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{first name} & \textbf{last name} & \textbf{email} & \textbf{country} & \textbf{organization}                & \textbf{corresponding?} \\ \hline
Geoffrey F.         & Schau              & schau@ohsu.edu & United States    & Oregon Health and Science University &  $\checkmark$                       \\ \hline
Shannon             & McWeeney           &                & United States    & Oregon Health and Science University &                         \\ \hline
Andrew              & Adey               & adey@ohsu.edu  & United States    & Oregon Health and Science University &  $\checkmark$                       \\ \hline
\end{tabular}%
}
}

